Analysis: The Sound of Teaching Music 1

### 2019-10-04 script/RNotebook/ioi.Rmd
add order analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@

@@

#### Plot
##### Bar graph
- Error bars show SEM of the mean
```{r ioi_dyn_var_all_bar, echo = FALSE}
p_ioi_dyn_var_all <- ggplot(data = ioi_dyn_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge()) +
  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
  theme(legend.position = "None")
p_ioi_dyn_var_all
```

@@

>>>>

- Order analysis (teaching/performing first?)
  + [All](#art_order_group_all)

@@

- Order analysis (teaching/performing first?)
  + [All](#dyn_order_group_all)

@@

```{r ioi_dyn_var_all_bar, echo = FALSE}
p_ioi_dyn_var_all <- ggplot(data = ioi_dyn_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge()) +
  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
  theme(legend.position = "None")
p_ioi_dyn_var_all
```

@@

## IOI-Order-Articulation
- Investigate whether there is any order effect of the conditions.

### 1. All{#art_order_group_all}
```{r ioi_art_order_all, echo = FALSE}
# For each participant
ioi_art_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_art,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
ioi_art_order <- cbind(ioi_art_order[,1:4], as.data.frame(ioi_art_order[,5]))
# Change colnames
colnames(ioi_art_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")

# Group mean
ioi_art_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_art_order,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_order_all <- cbind(ioi_art_order_all[,1:3], as.data.frame(ioi_art_order_all[,4]))
# Change colnames
colnames(ioi_art_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
print(ioi_art_order_all)
```

### Plot
#### Bar graph
- Error bars show SEM of the mean

```{r ioi_art_order_all_bar, echo = FALSE}
p_ioi_art_order_all <- ggplot(data = ioi_art_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge()) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(Order ~ .) +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_art_order_all
```

##### Box plot
```{r ioi_art_order_all_box, echo = FALSE}
p_ioi_art_order_all_box <- ggplot(data = ioi_art_order, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(Order ~ .) +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") +
  theme(legend.position = "None")
p_ioi_art_order_all_box
```

#### Statistics
##### Normality check
```{r normality7, echo = FALSE}
# Normality check
ioi_art_order_norm <- by(ioi_art_order$Mean, list(ioi_art_order$Condition, ioi_art_order$Order), shapiro.test)
print(ioi_art_order_norm)
```

##### Two-way ANOVA (ez)
```{r stats13, echo = FALSE}
ioi_art_order_aov <- ezANOVA(
  data = df_ioi_art
  , dv = .(IOI)
  , wid = .(SubNr)
  , between = .(Order)
  , within = .(Condition)
  , type = 3
  , detailed = TRUE
)
ioi_art_order_aov$ANOVA$F <- round(ioi_art_order_aov$ANOVA$F, 4)
ioi_art_order_aov$ANOVA$p <- round(ioi_art_order_aov$ANOVA$p, 4)
print(ioi_art_order_aov)
```

##### Two-way ANOVA (aov)
```{r stats14, echo = FALSE}
ioi_art_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_art_order)
print(summary(ioi_art_order_aov_2))
```

## IOI-Order-Dynamics
- Investigate whether there is any order effect of the conditions.

### 1. All{#dyn_order_group_all}
```{r ioi_dyn_order_all, echo = FALSE}
# For each pdynicipant
ioi_dyn_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_dyn,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
ioi_dyn_order <- cbind(ioi_dyn_order[,1:4], as.data.frame(ioi_dyn_order[,5]))
# Change colnames
colnames(ioi_dyn_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")

# Group mean
ioi_dyn_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_dyn_order,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_dyn_order_all <- cbind(ioi_dyn_order_all[,1:3], as.data.frame(ioi_dyn_order_all[,4]))
# Change colnames
colnames(ioi_dyn_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
print(ioi_dyn_order_all)
```

### Plot
#### Bar graph
- Error bars show SEM of the mean

```{r ioi_dyn_order_all_bar, echo = FALSE}
p_ioi_dyn_order_all <- ggplot(data = ioi_dyn_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge()) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(Order ~ .) +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_dyn_order_all
```

##### Box plot
```{r ioi_dyn_order_all_box, echo = FALSE}
p_ioi_dyn_order_all_box <- ggplot(data = ioi_dyn_order, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(Order ~ .) +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") +
  theme(legend.position = "None")
p_ioi_dyn_order_all_box
```

#### Statistics
##### Normality check
```{r normality8, echo = FALSE}
# Normality check
ioi_dyn_order_norm <- by(ioi_dyn_order$Mean, list(ioi_dyn_order$Condition, ioi_dyn_order$Order), shapiro.test)
print(ioi_dyn_order_norm)
```

##### Two-way ANOVA (ez)
```{r stats15, echo = FALSE}
ioi_dyn_order_aov <- ezANOVA(
  data = df_ioi_dyn
  , dv = .(IOI)
  , wid = .(SubNr)
  , between = .(Order)
  , within = .(Condition)
  , type = 3
  , detailed = TRUE
)
ioi_dyn_order_aov$ANOVA$F <- round(ioi_dyn_order_aov$ANOVA$F, 4)
ioi_dyn_order_aov$ANOVA$p <- round(ioi_dyn_order_aov$ANOVA$p, 4)
print(ioi_dyn_order_aov)
```

##### Two-way ANOVA (aov)
```{r stats16, echo = FALSE}
ioi_dyn_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_dyn_order)
print(summary(ioi_dyn_order_aov_2))
```
